Comparison of MCMC algorithms in Stochastic Volatility Models

### 2023-10-03 manuscript/main.tex
discussion section

    \subsection{Reweigthing rank statistics of Gaussian Mixture}
    \citet{kim1998stochastic} correct for approximation error in their method by using importance sampling weights. The reweighting procedure ensures that samples are drawn from the correct posterior density. This reweighting step is applied to the calculation of the expected value of the posterior density to get a weighted mean estimate from the samples drawn from the Gaussian mixture approximation. The rank statistic can be written as a function of the expected value, so the reweighting step can be applied to the calculation of the rank statistics to see if this correction improves the calibration of the sampling approach.

@@

    \subsection{Evaluating all models and parameters}
    As discussed, it is not practical to visually inspect the histograms of all parameters in the stochastic volatility model. Instead, the Chi squared statistic can be used to summarise the shape of the distribution and if desired, be used for hypothesis testing. However, isntead of performing hypothesis tests, another visual approach is taken to compare the calibration of the algorithms and parameters. 
    % However, performing all these tests suffers from the same overcommunication as presenting all the histograms at once.

@@

\section{Discussion}

>>>>

    \subsection{Reweigthing rank statistics of Gaussian Mixture}
    \citet{kim1998stochastic} correct for approximation error in their method by using importance sampling weights. The reweighting procedure ensures that samples are drawn from the correct posterior density. This is applied to the calculation of the expected value of the posterior density to get a weighted mean estimate from the samples drawn from the Gaussian mixture approximation. The rank statistic can be written as a function of the expected value, so the reweighting step can be applied to the calculation of the rank statistics to see if this correction improves the calibration of the sampling approach.

@@

    \subsection{Evaluating all models and parameters}
    As discussed, it is not practical to visually inspect the histograms of all parameters in the stochastic volatility model. Instead, the Chi squared statistic can be used to summarise the shape of the distribution and if desired, be used for hypothesis testing. However, instead of performing hypothesis tests, another visual approach is taken to compare the calibration of the algorithms and parameters. 
    % However, performing all these tests suffers from the same overcommunication as presenting all the histograms at once.

@@

\section{Discussion}
Overall, Hamiltonian Monte Carlo applied on the reparameterised model performed best based on the calibration and efficiency evaluation criteria. Centered parameterisation using HMC performed moderately well but struggled to produce calibrated posterior estimates for $\sigma^2$ and $\phi$ as well as generate independent samples for these parameters. Gaussian mixture approximation struggled to return calibrated posterior estimates and efficient samples for both parameteristions. However, the centered parameteristion results may improve if we increased the length of the Markov chain. This is discussed later in the limitations section.

These results demonstrate that the performance of a MCMC algorithm is sensitive to the paramterisation of the model in the context of calibration and efficiency. Furthermore, the paramterisation of a model is conditional on the choice of MCMC algorithm. Evaluating the effectiveness of a modelling strategy (i.e choice of model, paramterisation and sampler) on simulated data gives a lot of a priori information about how the model will perform before it is fit on real data. Performing SBC may help isolate confounding factors if any problems occur.

The priors chosen for this study came directly from \citet{kim1998stochastic}. The SBC results for the Gaussian mixture may improve if the priors were more informative - although this is only conjecture. Whether or not to use tighter priors depends on the modelling context and the problem at hand. If the model is expected to handle datasets generated by a set of priors, then uncalibrated results from SBC raises questions about the suitability about the model choice, MCMC or parameterisation.

Other parameterisaions of the stochastic volatility model are available as outlined in \citet{strickland2008parameterisation}, namely non centered in scale or both non centered in scale and location. The failure of non centered parameterisaion in location for the Gaussian mixture model could be due to a variety of reasons. This may be due to a particular software implementation when it comes to the Kalman Filter and Simulation Smoother (in which a specific software package was applied) that was not designed to handle the non centered parameterisation. Or perhaps this particular parameterisation doesn't suit the sampling strategy applied by KSC and another approach (e.g. HMC) might be more successful. Lots of different design and implementation details could be explored further. Pinpointing the reason why this model failed as well as exploring other configurations and MCMC algorithms in this context is left to future research.

% (other implementations exist as well but left for future research. lots of different design and implementation details could be explored. Let alone choice of algos.)
